---
title: "Homework-2024.11.04"
author: "By SA24204148"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Homework-2024.11.04}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Question 11.3

(a) Write a function to compute the kth term in

$$
\sum_{k=0}^{\infty} \frac{(-1)^k}{k!2^k} \frac{\|a\|^{2 k+2}}{(2 k+1)(2 k+2)} \frac{\Gamma\left(\frac{d+1}{2}\right) \Gamma\left(k+\frac{3}{2}\right)}{\Gamma\left(k+\frac{d}{2}+1\right)}
$$
where $d \ge 1$ is an integer, $a$ is a vector in $\mathbb{R}^d$, and $||\cdot||$ denotes the Euclidean norm. Perform the arithmetic so that the coefficients can be computed for (almost) arbitrarily large k and d. (This sum converges for all $a \in \mathbb{R}^d$).

(b) Modify the function so that it computes and returns the sum.

(c) Evaluate the sum when $a = (1, 2)^T$ .

## Answer

### (a)

The k-th term formula is:

$$
\text{term}_k = \frac{(-1)^k}{k!2^k} \frac{\|a\|^{2 k+2}}{(2 k+1)(2 k+2)} \frac{\Gamma\left(\frac{d+1}{2}\right) \Gamma\left(k+\frac{3}{2}\right)}{\Gamma\left(k+\frac{d}{2}+1\right)}.
$$

To avoid computing a ratio of two very large or very small numbers, the above formula is equivalent to following statement:

$$
term_k = (-1)^k\exp\{(2k+2)\log(||a||)-\log\Gamma(k+2)-(k+2)\log2+\log\Gamma\left(\frac{d+1}{2}\right)+\log\Gamma\left(k+\frac{1}{2}\right)-log\Gamma\left(k+\frac{d}{2}+1\right)\}
$$

```{r}
# Function to calculate the k-th term in the series
compute_kth_term <- function(k, a, d) {
  # Euclidean norm of vector a
  norm_a <- sqrt(sum(a^2))
  
  # Calculate each component of the exponent in the term
  term <- (-1)^k * exp(
    (2 * k + 2) * log(norm_a) - 
    lgamma(k + 2) - 
    (k + 2) * log(2) + 
    lgamma((d + 1) / 2) + 
    lgamma(k + 0.5) - 
    lgamma(k + d / 2 + 1)
  )
  
  return(term)
}
```

Alternatively, since $\Gamma(x+1)=x\Gamma(x)$, the ratio of the k+1th term to the kth term is $\frac{term_{k+1}}{term_k}=\frac{-(2k+1)||a||^2}{4(k+2)(k+d/2+1)}$. Using this equation, the k+1th term can be obtained by multiplying the ratio by the kth term which avoids computational issue.

### (b)

To compute the full sum, we'll loop over k until the terms are small enough to meet a convergence criterion.

```{r}
compute_sum <- function(a, d, k_out = 0 ,tolerance = 1e-10, max_iter = 1000) {
  total_sum <- 0
  k <- 0
  
  repeat {
    term <- compute_kth_term(k, a, d)
    total_sum <- total_sum + term
    
    # Check for convergence
    if (abs(term) < tolerance || k >= max_iter) {
      break
    }
    
    k <- k + 1
  }
  if(k_out) cat("k=", k, "\n")
  return(total_sum)
}
```

In R, the vectorized version appears to be faster than the method above using loops. Here is the function that computes the sum of the first k (enough large) terms using the vectorization method.

```{r}
compute_sum_vector <- function(a, d, k) {
  vec <- seq(k)-1
  a_norm <- sqrt(sum(a^2))
  i <- rep(c(1,-1), length = k)
  
  part1 <- lgamma(vec+1.5)-lgamma(vec+d/2+1)
  part2 <- (2*vec+2)*log(a_norm) - log(factorial(vec)) - vec*log(2) - log(2*vec+1)-log(2*vec+2)
  
  return(sum(i*exp(part1+part2))*gamma(d/2+0.5))
}
```

### (c)

```{r}
# Vector a and dimension d
a <- c(1, 2)
d <- 2

# Compute the sum
result <- compute_sum(a, d, k_out = 1)
print(paste("The sum is:", result))

compute_sum_vector(a, d, 17)
```
For the method using loop, convergence is reached when k = 17 and the result of the calculation is 1.53216399886018, which is almost identical to the result using the vectorization method(1.532164).

```{r}
d = 10
a = seq(d)

system.time(
  {
    for (i in seq(1000)) {
      compute_sum(a, d)
    }
  }
)

system.time(
  {
    for (i in seq(1000)) {
      compute_sum_vector(a, d, 17)
    }
  }
)
```

When d = 10, repeating both of these methods 10,00 times, it can be seen that the vectorized method is much faster than the loop method.

## Question 11.4

Find the intersection points $A(k)$ in $(0,\sqrt{k})$ of the curves
 
$$
 S_{k-1}(a)=P\left(t(k-1)>\sqrt{\frac{a^2(k-1)}{k-a^2}}\right)
$$

and

$$
S_k(a)=P\left(t(k)>\sqrt{\frac{a^2 k}{k+1-a^2}}\right)
$$

for $k = 4 : 25, 100, 500, 1000$, where $t(k)$ is a Student $t$ random variable with $k$ degrees of freedom. 

## Answer

 + Determine the Intersection Point: The intersection point $A(k)$ is the value $a$ at which the two curves $S_{k-1}(a)$ and $S_{k}(a)$ are equal.
 + Define the Equation: Substitute the expressions for $S_{k-1}(a)$ and $S_{k}(a)$into an equation, and solve for $a$ numerically. 
 + Numerical Solution: Since this is an equation, we can use a numerical solver in R, such as uniroot, to find the value of $a$ that meets this condition.

```{r}
# Define a function to calculate Sk-1(a) - Sk(a)
compute_difference <- function(a, k) {
  # Calculate Sk-1(a)
  critical_value_k_minus_1 <- sqrt(a^2 * (k - 1) / (k - a^2))
  prob_k_minus_1 <- pt(critical_value_k_minus_1, df = k - 1, lower.tail = FALSE)
  
  # Calculate Sk(a)
  critical_value_k <- sqrt(a^2 * k / (k + 1 - a^2))
  prob_k <- pt(critical_value_k, df = k, lower.tail = FALSE)
  
  # Return the difference
  return(prob_k_minus_1 - prob_k)
}

# Set the value of k, for example, k = 4
k_set <- c(4 : 25, 100, 500, 1000)

result1 <- NULL

for (k in k_set) {
  # Use uniroot to find the value of a in the interval (0, sqrt(k))
  result <- uniroot(compute_difference, interval = c(1e-5, ifelse(k>100,sqrt(k)-4,sqrt(k)-1e-5)), k = k)
  a_intersection <- result$root
  result1 <- c(result1, a_intersection)

  cat("Intersection point A(k) for k =", k, "is a =", a_intersection, "\n")
}
```

Note that to avoid computational problems, the interval used in the actual uniroot function shrinks inward $\epsilon$

## Question 11.5

Write a function to solve the equation

$$
\begin{aligned}
& \frac{2 \Gamma\left(\frac{k}{2}\right)}{\sqrt{\pi(k-1)} \Gamma\left(\frac{k-1}{2}\right)} \int_0^{c_{k-1}}\left(1+\frac{u^2}{k-1}\right)^{-k / 2} d u \\
& \quad=\frac{2 \Gamma\left(\frac{k+1}{2}\right)}{\sqrt{\pi k} \Gamma\left(\frac{k}{2}\right)} \int_0^{c_k}\left(1+\frac{u^2}{k}\right)^{-(k+1) / 2} d u
\end{aligned}
$$
for $a$, where

$$
c_k=\sqrt{\frac{a^2 k}{k+1-a^2}} .
$$

Compare the solutions with the points $A(k)$ in Exercise 11.4.

## Answer

Use the integrate function to calculate the integrals of the left and right sides of the equation, and to avoid calculating the ratio of two larger numbers, use the $\exp(\log x)=x$ method.

```{r}
# Define the function to compute the integrals and find 'a'
solve_for_a <- function(k) {
  # Define the target function based on the equation
  target_function <- function(a) {
    # Compute ck and ck-1
    ck <- sqrt(a^2 * k / (k + 1 - a^2))
    ck_minus_1 <- sqrt(a^2 * (k - 1) / (k - a^2))
    
    # Compute the left-hand side integral
    left_integral <- integrate(function(u) (1 + u^2 / (k - 1))^(-k / 2), 0, ck_minus_1)$value
    left_side <- 2 * exp(lgamma(k / 2) - lgamma((k - 1) / 2)) / (sqrt( (k - 1))) * left_integral
    
    # Compute the right-hand side integral
    right_integral <- integrate(function(u) (1 + u^2 / k)^(-(k + 1) / 2), 0, ck)$value
    right_side <- 2 * exp(lgamma((k + 1) / 2)-lgamma(k / 2)) / (sqrt( k) )  * right_integral
    
    # Return the difference
    return(left_side - right_side)
  }
  
  # Use a root-finding method to find 'a' such that target_function(a) = 0
  result <- uniroot(target_function, c(1, 1.8))$root
  return(result)
  
}

result2 <- NULL

# Example usage with a specific value of k
k_set <- c(4 : 25, 100, 500, 1000)
for (k in k_set) {
  a_solution <- solve_for_a(k)
  result2 <- c(result2, a_solution)
  cat("the root for k =",k , "is", a_solution,"\n")
}
```

As a special note, due to the computational issue, the interval is specifically limited to (1,1.8) when solving for the zeros here.

Comparison with the results of Exercise 11.4 shows that they are almost identical.

```{r}
result1-result2
```


## Question 

Suppose $T_1, \cdots , T_n$ are i.i.d. samples drawn from the exponential distribution with expectation $\lambda$. Those values greater than $\tau$ are not observed due to right censorship, so that the observed values are $Y_i=T_i \mathbb{I}\left(T_i \leq \tau\right)+\tau \mathbb{I}\left(T_i>\tau\right)$, $i = 1,\cdots, n$. Suppose $\tau= 1$ and the observed $Y_i$ values are as follows:

$$
0.54, 0.48, 0.33, 0.43, 1.00, 1.00, 0.91, 1.00, 0.21, 0.85
$$

Use the E-M algorithm to estimate $\lambda$, compare your result with the observed data MLE (note: $Y_i$ follows a mixture distribution).

## Answer

First use the MLE method to estimate the $\lambda$.

Due to $Y_i=T_i \mathbb{I}\left(T_i \leq \tau\right)+\tau \mathbb{I}\left(T_i>\tau\right)$,the distribution function of $Y_i$ is given by 

$$
\begin{aligned}
P(Y_i\leq t)&=  \mathbb{I}(\tau\leq t) + \mathbb{I}(\tau>t)P(T_i\leq t) \\
            &=  \mathbb{I}(\tau\leq t) + \mathbb{I}(\tau>t)(1-e^{-\lambda t})
\end{aligned}
$$

Therefore, the density function of $Y_i$ is

$$
f_{Y_i}(t)=
\begin{cases}
\lambda e^{-\lambda t}, & t<\tau\\
e^{-\lambda \tau}, & t=\tau
\end{cases}
$$

By minimizing the following log-likelihood function, the estimate of lambda is obtained as $\hat{\lambda}_{mle}=\frac{\sum_{i=1}^n \mathbb{I}(Y_i<\tau)}{\sum_{i=1}^n Y_i}$

$$
l(\lambda) = -\lambda\sum_{i=1}^n Y_i +\sum_{i=1}^n \mathbb{I}(Y_i<\tau)\log\lambda
$$

```{r}
y = c(0.54, 0.48, 0.33, 0.43, 1.00, 1.00, 0.91, 1.00, 0.21, 0.85)
tau = 1
lambda1 = sum(y<tau)/sum(y)
cat("The estimate of lambda by mle is", lambda1, "\n")
```
Next use the E-M algorithm.

 + Complete data: $X = (Y_i, T_i)$
 + Observed data: $Y_i$
 + missing data: $T_i$ 
 + likelihood function: $L(\lambda | Y_i,T_i) = \Pi_{i=1}^n \lambda e^{-\lambda T_i}$

E-step: 

Given the observed data $Y_i$, the distribution function for $T_i$ is 

$$
\begin{aligned}
P(T_i\leq t|Y_i)&=\mathbb{I}(t<\tau)\mathbb{I}(Y_i\leq t) + \mathbb{I}(t\ge\tau)(\mathbb{I}(Y_i<\tau)+\mathbb{I}(Y_i=\tau)P(T_i\leq t|T_i\ge\tau))\\
&= \mathbb{I}(t<\tau)\mathbb{I}(Y_i\leq t) + \mathbb{I}(t\ge\tau)(\mathbb{I}                                                                                                                                                                                                            (Y_i<\tau)+\mathbb{I}(Y_i=\tau)\frac{e^{-\lambda\tau}-e^{-\lambda t}}{e^{-\lambda\tau}})\\  
\end{aligned}
$$

Therefore, the density function of $T_i|Y_i$ is

$$
f_{T_i|Y_i}(t)=
\begin{cases}
1, & t = Y_i\\ 
\lambda e^{\lambda(\tau-t)}, & t\ge\tau, Y_i=\tau
\end{cases}
$$

$$
\mathbb{E}(T_i|Y_i)=\mathbb{I}(Y_i\not=\tau)Y_i+\mathbb{I}(Y_i=\tau)(\tau+\frac{1}{\lambda})
$$

M-step:

Given $T_i=\mathbb{E}_{\lambda_k}(T_i|Y_i)$, maximize the log-likelihood to obtain an estimate of $\lambda_{k+1}$ as $\frac{n}{T_i}$:

$$
l(\lambda | Y_i,T_i)=-\lambda\sum T_i+n\log\lambda
$$

```{r}
# E-M algorithm
tol <- 1e-6  # Convergence tolerance
max_iter <- 1000  # Maximum number of iterations
lambda = lambda1
for (iter in 1:max_iter) {
  # E-step: Calculate expected T_i for censored and uncensored values
  expected_T <- ifelse(y < tau, y, tau + 1 / lambda)
  
  # M-step: Update lambda
  new_lambda <- length(y) / sum(expected_T)
  
  # Check convergence
  if (abs(new_lambda - lambda) < tol) {
    lambda <- new_lambda
    break
  }
  
  # Update lambda for the next iteration
  lambda <- new_lambda
}

# Output the estimated lambda
lambda
```

The estimate of $\lambda$ by E-M algorithm is 1.037037, which is almost identical to the estimate of $\lambda$ by mle(1.037037). 
